---
title: "ld_analysis"
author: "sq-96"
date: "2021-07-05"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

## Objective
Comparing eigenvalues and eigenvectors of LD matrices calcualted with different subsamples.

## Method
Data is from 1000 Genome EUR ancestry (~500 individuals). Samples are randomly splitted into 3 subsets, each has 160 individuals. Using PLINL to alculate LD matrices with variants (MAF > 0.05) in three SCZ loci. 

## Result
```{r,echo=FALSE}
library(plot.matrix)
library(RColorBrewer)
```

```{r,message=FALSE,results=FALSE,echo=FALSE,eval=FALSE}
set.seed(2021)
sample_list <- read.table("/project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.1.fam",header=FALSE,sep=" ")
sample_list <- sample_list[sample(1:489,480),c("V1","V2")]
sample_list_split <- split(sample_list, sample(rep(1:3,times=c(160,160,160))))
subsample1 <- sample_list_split$`1`
subsample2 <- sample_list_split$`2`
subsample3 <- sample_list_split$`3`

write.table(subsample1,file = "./data/EUR_subpop1.txt",sep='\t',quote = FALSE,col.names = FALSE,row.names = FALSE)
write.table(subsample2,file = "./data/EUR_subpop2.txt",sep='\t',quote = FALSE,col.names = FALSE,row.names = FALSE)
write.table(subsample3,file = "./data/EUR_subpop3.txt",sep='\t',quote = FALSE,col.names = FALSE,row.names = FALS```
```

```{r,echo=FALSE}
cor_cal <- function(vec1,vec2,num=10){
  cor_vec <- c()
  for(i in 1:num){
    cor_vec[i] <- round(cor(vec1[,i],vec2[,i]),3)
  }
  return(cor_vec)
}

cor_matrix_cal <- function(vec1,vec2){
  num <- dim(vec1)[2]
  cor_matrix <- matrix(NA,nrow = num,ncol = num)
  for(i in 1:num){
    for(j in 1:num){
      cor_matrix[i,j] <- round(cor(vec1[,i],vec2[,j]),3)
    }
  }
  return(cor_matrix)
}
```

### chr15 91416560-91429040
```{r,message=FALSE,results=FALSE,echo=FALSE,eval=FALSE}
system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.15 --keep ./data/EUR_subpop1.txt --chr 15 --from-bp 91416560 --to-bp 91429040 --write-snplist --maf 0.05 --r --matrix --out ./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop1")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.15 --keep ./data/EUR_subpop2.txt --chr 15 --from-bp 91416560 --to-bp 91429040 --write-snplist --maf 0.05 --r --matrix --out ./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop2")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.15 --keep ./data/EUR_subpop3.txt --chr 15 --from-bp 91416560 --to-bp 91429040 --write-snplist --maf 0.05 --r --matrix --out ./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop3")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.15 --chr 15 --from-bp 91416560 --to-bp 91429040 --write-snplist --maf 0.05 --r --matrix --out ./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR")
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
ld1 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop1.ld',header=FALSE)
snp_list1 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop1.snplist',header=FALSE)
colnames(ld1) <- snp_list1$V1
rownames(ld1) <- snp_list1$V1

ld2 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop2.ld',header=FALSE)
snp_list2 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop2.snplist',header=FALSE)
colnames(ld2) <- snp_list2$V1
rownames(ld2) <- snp_list2$V1

ld3 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop3.ld',header=FALSE)
snp_list3 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR_subpop3.snplist',header=FALSE)
colnames(ld3) <- snp_list3$V1
rownames(ld3) <- snp_list3$V1

ld4 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR.ld',header=FALSE)
snp_list4 <- read.table('./data/chr15_91416560_91429040_EUR/chr15_91416560_91429040_EUR.snplist',header=FALSE)
colnames(ld4) <- snp_list4$V1
rownames(ld4) <- snp_list4$V1
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
index <- Reduce(intersect, list(snp_list1$V1,snp_list2$V1,snp_list3$V1,snp_list4$V1))
ld1 <- ld1[index,index]
ld2 <- ld2[index,index]
ld3 <- ld3[index,index]
ld4 <- ld4[index,index]

result1 <- eigen(ld1)
result2 <- eigen(ld2)
result3 <- eigen(ld3)
result4 <- eigen(ld4)

value1 <- round(result1$values,3)
value2 <- round(result2$values,3)
value3 <- round(result3$values,3)
value4 <- round(result4$values,3)

vector1 <- result1$vectors
vector2 <- result2$vectors
vector3 <- result3$vectors
vector4 <- result4$vectors
```

#### The first ten eigenvalues
```{r,echo=FALSE}
value_df <- rbind(value1[1:10],value2[1:10],value3[1:10],value4[1:10])
rownames(value_df) <- c("EUR_sub1","EUR_sub2","EUR_sub3","EUR_All")
value_df
```

#### Pairwise correlation of the ith eigenvectors
```{r,echo=FALSE}
vector_cor_df <-rbind(cor_cal(vector1,vector2),cor_cal(vector1,vector3),cor_cal(vector2,vector3),
                      cor_cal(vector1,vector4),cor_cal(vector2,vector4),cor_cal(vector3,vector4))
rownames(vector_cor_df) <- c("sub1.vs.sub2","sub1.vs.sub3","sub2.vs.sub3",
                             "sub1.vs.All","sub2.vs.All","sub3.vs.All")
vector_cor_df
```

#### Pairwise correlation between two sets of eigenvectors
```{r,fig.height=5, fig.width=6,echo=FALSE}
matrix_cor_df <- cor_matrix_cal(vector1,vector3)
par(mar=c(5, 4, 4, 4))
plot(as.matrix(matrix_cor_df), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='sub1 vs sub2', col=brewer.pal(5,name = "RdBu"),spacing.key=c(2,2,1))
```

#### Heatmap of LD matrix
```{r,fig.height=15, fig.width=15,echo=FALSE}
par(mar=c(5, 4, 4, 2),mfrow=c(2,2))
plot(as.matrix(ld1), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub1', col=rev(brewer.pal(8,name = "RdYlBu")),spacing.key=c(2,2,1))
plot(as.matrix(ld2), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub2', col=rev(brewer.pal(8,name = "RdYlBu")),spacing.key=c(2,2,1))
plot(as.matrix(ld3), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub3', col=rev(brewer.pal(8,name = "RdYlBu")),spacing.key=c(2,2,1))
plot(as.matrix(ld4), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_All',  col=rev(brewer.pal(8,name = "RdYlBu")),spacing.key=c(2,2,1))
```

### chr8 4177794-4192544
```{r,message=FALSE,results=FALSE,echo=FALSE,eval=FALSE}
system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.8 --keep ./data/EUR_subpop1.txt --chr 8 --from-bp 4177794 --to-bp 4192544 --write-snplist --maf 0.05 --r --matrix --out ./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop1")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.8 --keep ./data/EUR_subpop2.txt --chr 8 --from-bp 4177794 --to-bp 4192544 --write-snplist --maf 0.05 --r --matrix --out ./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop2")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.8 --keep ./data/EUR_subpop3.txt --chr 8 --from-bp 4177794 --to-bp 4192544 --write-snplist --maf 0.05 --r --matrix --out ./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop3")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.8 --chr 8 --from-bp 4177794 --to-bp 4192544 --write-snplist --maf 0.05 --r --matrix --out ./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR")
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
ld1 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop1.ld',header=FALSE)
snp_list1 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop1.snplist',header=FALSE)
colnames(ld1) <- snp_list1$V1
rownames(ld1) <- snp_list1$V1

ld2 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop2.ld',header=FALSE)
snp_list2 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop2.snplist',header=FALSE)
colnames(ld2) <- snp_list2$V1
rownames(ld2) <- snp_list2$V1

ld3 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop3.ld',header=FALSE)
snp_list3 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR_subpop3.snplist',header=FALSE)
colnames(ld3) <- snp_list3$V1
rownames(ld3) <- snp_list3$V1

ld4 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR.ld',header=FALSE)
snp_list4 <- read.table('./data/chr8_4177794_4192544_EUR/chr8_4177794_4192544_EUR.snplist',header=FALSE)
colnames(ld4) <- snp_list4$V1
rownames(ld4) <- snp_list4$V1
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
index <- Reduce(intersect, list(snp_list1$V1,snp_list2$V1,snp_list3$V1,snp_list4$V1))
ld1 <- ld1[index,index]
ld2 <- ld2[index,index]
ld3 <- ld3[index,index]
ld4 <- ld4[index,index]

result1 <- eigen(ld1)
result2 <- eigen(ld2)
result3 <- eigen(ld3)
result4 <- eigen(ld4)

value1 <- round(result1$values,3)
value2 <- round(result2$values,3)
value3 <- round(result3$values,3)
value4 <- round(result4$values,3)

vector1 <- result1$vectors
vector2 <- result2$vectors
vector3 <- result3$vectors
vector4 <- result4$vectors
```

#### The first ten eigenvalues
```{r,echo=FALSE}
value_df <- rbind(value1[1:10],value2[1:10],value3[1:10],value4[1:10])
rownames(value_df) <- c("EUR_sub1","EUR_sub2","EUR_sub3","EUR_All")
value_df
```

#### Pairwise correlation of the ith eigenvectors
```{r,echo=FALSE}
vector_cor_df <-rbind(cor_cal(vector1,vector2),cor_cal(vector1,vector3),cor_cal(vector2,vector3),
                      cor_cal(vector1,vector4),cor_cal(vector2,vector4),cor_cal(vector3,vector4))
rownames(vector_cor_df) <- c("sub1.vs.sub2","sub1.vs.sub3","sub2.vs.sub3",
                             "sub1.vs.All","sub2.vs.All","sub3.vs.All")
vector_cor_df
```

#### Pairwise correlation between two sets of eigenvectors
```{r,fig.height=5, fig.width=6,echo=FALSE}
matrix_cor_df <- cor_matrix_cal(vector1[,1:30],vector2[,1:30])
par(mar=c(5, 4, 4, 4))
plot(as.matrix(matrix_cor_df), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='sub1 vs sub2', col=brewer.pal(5,name = "RdBu"),spacing.key=c(2,2,1))
```

#### Heatmap of LD matrix
```{r,fig.height=10, fig.width=10,echo=FALSE,eval=FALSE}
par(mar=c(5, 4, 4, 2),mfrow=c(2,2))
plot(as.matrix(ld1), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub1', col=brewer.pal(8,name = "Reds"),spacing.key=c(2,2,1))
plot(as.matrix(ld2), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub2', col=brewer.pal(8,name = "Reds"),spacing.key=c(2,2,1))
plot(as.matrix(ld3), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_sub3', col=brewer.pal(8,name = "Reds"),spacing.key=c(2,2,1))
plot(as.matrix(ld4), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='EUR_All',  col=brewer.pal(8,name = "Reds"),spacing.key=c(2,2,1))
```

### chr10 18681005-18770105
```{r,message=FALSE,results=FALSE,echo=FALSE,eval=FALSE}
system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.10 --keep ./data/EUR_subpop1.txt --chr 10 --from-bp 18681005 --to-bp 18770105 --write-snplist --maf 0.05 --r --matrix --out ./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop1")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.10 --keep ./data/EUR_subpop2.txt --chr 10 --from-bp 18681005 --to-bp 18770105 --write-snplist --maf 0.05 --r --matrix --out ./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop2")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.10 --keep ./data/EUR_subpop3.txt --chr 10 --from-bp 18681005 --to-bp 18770105 --write-snplist --maf 0.05 --r --matrix --out ./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop3")

system("/software/plink-1.90-el7-x86_64/plink --bfile /project2/xinhe/1kg/1000G_EUR_Phase3_plink/1000G.EUR.QC.10 --chr 10 --from-bp 18681005 --to-bp 18770105 --write-snplist --maf 0.05 --r --matrix --out ./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR")
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
ld1 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop1.ld',header=FALSE)
snp_list1 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop1.snplist',header=FALSE)
colnames(ld1) <- snp_list1$V1
rownames(ld1) <- snp_list1$V1

ld2 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop2.ld',header=FALSE)
snp_list2 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop2.snplist',header=FALSE)
colnames(ld2) <- snp_list2$V1
rownames(ld2) <- snp_list2$V1

ld3 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop3.ld',header=FALSE)
snp_list3 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR_subpop3.snplist',header=FALSE)
colnames(ld3) <- snp_list3$V1
rownames(ld3) <- snp_list3$V1

ld4 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR.ld',header=FALSE)
snp_list4 <- read.table('./data/chr10_18681005_18770105_EUR/chr10_18681005_18770105_EUR.snplist',header=FALSE)
colnames(ld4) <- snp_list4$V1
rownames(ld4) <- snp_list4$V1
```

```{r,message=FALSE,results=FALSE,echo=FALSE}
index <- Reduce(intersect, list(snp_list1$V1,snp_list2$V1,snp_list3$V1,snp_list4$V1))
ld1 <- ld1[index,index]
ld2 <- ld2[index,index]
ld3 <- ld3[index,index]
ld4 <- ld4[index,index]

result1 <- eigen(ld1)
result2 <- eigen(ld2)
result3 <- eigen(ld3)
result4 <- eigen(ld4)

value1 <- round(result1$values,3)
value2 <- round(result2$values,3)
value3 <- round(result3$values,3)
value4 <- round(result4$values,3)

vector1 <- result1$vectors
vector2 <- result2$vectors
vector3 <- result3$vectors
vector4 <- result4$vectors
```

#### The first ten eigenvalues
```{r,echo=FALSE}
value_df <- rbind(value1[1:10],value2[1:10],value3[1:10],value4[1:10])
rownames(value_df) <- c("EUR_sub1","EUR_sub2","EUR_sub3","EUR_All")
value_df
```

#### Pairwise correlation of the ith eigenvectors
```{r,echo=FALSE}
vector_cor_df <-rbind(cor_cal(vector1,vector2),cor_cal(vector1,vector3),cor_cal(vector2,vector3),
                      cor_cal(vector1,vector4),cor_cal(vector2,vector4),cor_cal(vector3,vector4))
rownames(vector_cor_df) <- c("sub1.vs.sub2","sub1.vs.sub3","sub2.vs.sub3",
                             "sub1.vs.All","sub2.vs.All","sub3.vs.All")
vector_cor_df
```

#### Pairwise correlation between two sets of eigenvectors
```{r,fig.height=5, fig.width=6,echo=FALSE}
matrix_cor_df <- cor_matrix_cal(vector1[,1:30],vector2[,1:30])
par(mar=c(5, 4, 4, 4))
plot(as.matrix(matrix_cor_df), axis.col=NULL, axis.row=NULL, xlab='', ylab='', main='sub1 vs sub2', col=brewer.pal(5,name = "RdBu"),spacing.key=c(2,2,1))
```
